---
title: "Homework 4"
author: "Tina Zhang"
date: "2019/5/20"
output: pdf_document
---

```{r setup, message = FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
packages <- c("readr","proxy","dplyr","tidytext","ggplot2","SnowballC","topicmodels",
              "stm","tidyr","mixtools","tm")
load.packages <- function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
}
lapply(packages, load.packages)
setwd("C:/Users/tzwhi/Desktop/Northwestern/311-2")
```

#Question 1
```{r fig.height= 15, fig.width= 10}
#1
inmates <- read.csv("tx_deathrow_full.csv", stringsAsFactors = FALSE, encoding = "UTF-8") %>%
  mutate(doc_id = row_number(), text = Last.Statement) %>%
  select(doc_id, text, everything())
corp <- VCorpus(DataframeSource(inmates)) %>%
  tm_map(removePunctuation) %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removeWords, stopwords("english")) %>%
  tm_map(stemDocument)
dtm <- DocumentTermMatrix(corp)
empty_rows <- which(rowSums(as.matrix(dtm))==0)
dtm <- corp[-empty_rows] %>% DocumentTermMatrix()

#2
lda <- LDA(dtm, k=10, control = list(seed=25))

#3
tidy(lda) %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  mutate(term = reorder(term, beta)) %>%
  ggplot(aes(term, beta, fill=factor(topic))) + geom_col(show.legend = FALSE) +
  facet_wrap(~topic, scales = "free_y", nrow = 5) + coord_flip() +
  xlab("Term") + ylab("Beta") + 
  labs(title = "LDA Analysis of Texas Death Row Statements, K = 10",
                      subtitle = "Top 10 Most Likely Words by Topic")
```

#Question 2
```{r, results= "hide"}
#1
out <- stm::readCorpus(dtm, type = "slam")

#2
mod.out <- stm(documents = out$documents, vocab = out$vocab, K = 10,
               prevalence = ~Race, data = inmates[-empty_rows, ])
```

```{r}
#3
summary(mod.out)
```

4. The topics found when conditioning on venue appear more differentiated than those found using standard LDA. For example, topics 8 and 5 appear tied to thankfulness and love, while topics 2 and 6 appear tied to religion and forgiveness. 
In contrast, there's a lot of overlap in the top words for each topic in LDA, and it's harder to identify what each topic is about. For example, the word "love" appears in 9 of the 10 topics, while words like "god", "forgiv", and "sorri", and "thank" also appear in many of the topics.